---
title: "Plotting collections"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting collections}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=10, 
  fig.height=5
)
```

## Introduction

The `autoplot` method provides a simple plotting interface to 
easily produce plots of the spectra contained in a collection in their current 
processing status.

```{r setup}
library(tidyverse)
library(tidySpectR)
```

## Basic usage

First let us have a look at the `fa_nmr` dataset that is provided 
with this package:

```{r rawplot}
print(fa_nmr)

autoplot(fa_nmr) +
    scale_x_reverse() # Norm für NMR is a reversed x axis
```

This is a bit clutered so we can try to separate a bit the spectra.
For this we can play with two parameters: `offset_y` and `offset_x`
which will offset each spectra on their respective axes. 

The offset value must be provided in the axis unit!

```{r rawoffset}
autoplot(fa_nmr, offset_y = 1e+09) +
    theme(legend.position = "none") +
    scale_x_reverse()

autoplot(fa_nmr, offset_y = 1e+09, offset_x = 1) +
    theme(legend.position = "none") +
    scale_x_reverse()
```

If we are not interested in visualizing individual spectra we can 
simply plot the average supectrum of the label-average spectra.
In this case the average spectrum will be respresented by a 
solid line and the range of data by a transparent ribbon.

```{r averages}
autoplot(fa_nmr, type = "average") +
    theme(legend.position = "none") +
    scale_x_reverse()

autoplot(fa_nmr, type = "label_average", offset_y = 1e+09, offset_x = 1) +
    scale_x_reverse()
```

## Plotting specific regions

It is possible to look at a specific region by preprocessing
the collection with the `extract` method
Let's take a closer look at what is happening in the 3.3 - 3.9 ppm region:

```{r crop}
fa_nmr %>% 
    extract(from = 3.3, to = 3.9) %>%
    autoplot(offset_y = 2e+08) +
        theme(legend.position = "none") +
        scale_x_reverse()
```

## Controlling bucketting

Visualization is especially needed to control the bucketting process.
Let's try different bucketting methods and check their results 
on the raw spectra.

For this example we will focus only on the 3.3 - 3.9 ppm region.

```{r bucketting}
uniform <- fa_nmr %>% 
    extract(from = 3.3, to = 3.9) %>%
    bucket_uniform(width = 0.04) %>%
    pull_breaks()
    
fa_nmr %>% 
    extract(from = 3.3, to = 3.9) %>%
    autoplot(fa_nmr, type = "average") +
    scale_x_reverse() +
    geom_vline(xintercept = uniform, color = "red", alpha = 0.5) +
    theme(legend.position = "none") +
    ggtitle("Uniform bucketting 0.04ppm")
    
opti <- fa_nmr %>% 
    extract(from = 3.3, to = 3.9) %>%
    bucket_optimized(initial_width = 0.04, slackness = 0.25) %>%
    pull_breaks()
    
fa_nmr %>% 
    extract(from = 3.3, to = 3.9) %>%
    autoplot(fa_nmr, type = "average") +
    scale_x_reverse() +
    geom_vline(xintercept = opti, color = "red", alpha = 0.5) +
    theme(legend.position = "none") +
    ggtitle("Optimized bucketting 0.04ppm")
```

This could still be optimized but we can now visualize the 
processed data.

```{r processed}
fa_nmr %>% 
    extract(from = 3.3, to = 3.9) %>%
    bucket_optimized(initial_width = 0.04, slackness = 0.25) %>%
    autoplot(type = "average") +
    scale_x_reverse()
```





